A North Sámi to South Sámi machine translation prototype

### 2014-02-08 paper/paper.tex
A bit more about the languages.

There are also differences within the Saami family.

sme-sma.

An overview \cite{sammallahti98}.

>>>>

North and South Saami are not mutually intellegible, due both to
linguistic distance and to radically different prinicples for their
literary languages. An analogy would be the difference between English 
and Frisian.

Grammatical similarities between North and South Saami include the 
following: Both languages have 3 persons and numbers, they have a similar
system for postpositions and verb derivation. Noun phrase syntax is 
similar, and the case systems are almost identical.

Taking a closer look, there are still differences. In both North and
South Saami, negation is expressed with a negation verb which inflects
for person and number, but in 
South Saami, this verb is inflected for Tense as well. South Saami
has OV word order and lacks a copula in predicative constructions, 
whereas North Saami uses VO in simple sentences, and requires a copula.

Although the case systems are similar, North Saami has a locative case
that covers the semantic field of both inessive and elative in 
South Saami, the locative must thus be split into two cases. There is
also some differences in case usage, plural objects are accusative
in North Saami, but nominative for indefinite and accusative for 
definite objects in South Saami.

For a machine translation system the ortographic differences 
imply that apart from person
names and acronyms there will be no free rides during the conversion 
process. The core vocabulary is distinct, even recent loans from
the same donor languages are different, and the vocabulary coverage
for a working system must thus be very good.

An overview over the differences between the two languages can
be found in \cite{sammallahti98}.
